\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2021-08-28 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updates

\begin{figure}[t]
\centering
\includegraphics[width=.7\linewidth]{BlockDiagm.pdf}
\caption{\small{Schematic of the proposed proximal algorithm for propagating the joint state PDF as probability weighted scattered point cloud $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$. The states $\{\tilde{\bm{x}}_{k}^{i}\}_{i=1}^{N}$ are updated by the Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}); the corresponding probability weights $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are updated via discrete version of the proximal recursion (\ref{KineticProx}) as detailed in Sec. \ref{sec:ProxAlgo}.}}
\label{fig:BlockDiagm}
\vspace*{-0.1in}
\end{figure}

@@

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.95\linewidth]{ComputationalTimeSyntheticIEEE14busCase1and2combined.png}
\caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ for the simulation set up in Sec. \ref{subsec:NumSimIEEEBus}.}}
\vspace*{-0.1in}
\label{fig:CompTimeIEEE14bus}
\end{figure}

@@

\noindent\textbf{Case I.} {\red{Nominal case where ...}}

\noindent\textbf{Case II.} {\red{post-contingency simulation after the line 13 fails.}}

@@

For all the three cases, we used the randomly generated noise coefficients $\sigma_{i}\in [1,5]$ as
\[(\sigma_{1},\hdots,\sigma_5) = \left(2.4628, 4.9266, 4.8724, 1.4215, 3.8681\right).\]

We take the initial joint PDF as
\begin{align}
\rho_{0}&\equiv\rho(t=0,\bm{\theta}(0),\bm{\omega}(0)) = \vartheta_{0}\left(\bm{\theta}(0)\right)\Omega_{0}\left(\bm{\omega}(0)\right)\nonumber\\
&=\underbrace{\left(\prod_{i=1}^{n=5}\dfrac{\exp\left(\kappa_{i}\cos\left(2\theta_{i}(0)-\mu_{\theta_{i}(0)}\right)\right)}{2\pi I_{0}(\kappa_{i})}\right)}_{=:\vartheta_{0}\left(\bm{\theta}(0)\right)}\nonumber\\
	&\qquad\times \underbrace{{\rm{Unif}}\left([-0.1,0.1]\;\text{rad/s}\right)}_{=:\Omega_{0}\left(\bm{\omega}(0)\right)},
\label{rho0IEEE14}		
\end{align}
whose $\bm{\theta}$ marginal $ \vartheta_{0}\left(\bm{\theta}(0)\right)$ is a product von Mises PDF\footnote{the factor $2$ inside the cosine ensures the range $[0,2\pi)$; it can be dispensed if we instead use the range $[-\pi,\pi)$ for the angular variables, as common in the directional statistics literature \cite{mardia2009directional}.} \cite{mardia2008multivariate,mardia2014some} supported over $\mathbb{T}^{5}$ with mean angles (in rad)
\begin{align}
\left(\mu_{\theta_{1}}(0),\hdots,\mu_{\theta_{5}}(0)\right)=\left(0, 6.1963, 6.0612, 6.0350, 6.0500\right),
\label{MeanAngleInitialIEEE14}	
\end{align}
and concentration parameter $\kappa_{i}\geq 0$ set as 
\[(\kappa_{1},\hdots,\kappa_5) = \left(5,6,7,4,5\right).\]
In \eqref{rho0IEEE14}, $I_{0}(\cdot)$ denotes the modified Bessel function of the first kind with order zero, given by $I_{0}(\kappa) = \sum_{r=0}^{\infty} \frac{\left(\kappa^{2}/4\right)^{r}}{(r!)^{2}}$, and can be evaluated via MATLAB command \texttt{besseli(0,$\cdot$)}.

@@

Fig. \ref{fig:IEEE14omegamarginalsCase12} shows the corresponding univariate $\omega_{i}$-marginal PDFs. In particular, Fig. \ref{fig:IEEE14omegamarginalsCase12}(c) helps compare the $\omega_{4}$ marginal evolution for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios. Since Case II considers the failure of line 13, from Fig. \ref{fig:Motivating} we intuitively expect that the transient effects of this fault will be prominent for bus 6 (generator 4). Indeed, Fig. \ref{fig:IEEE14omegamarginalsCase12}(c) shows that although the $\omega_{4}$ marginals in both Case I and II, at all times have highest probability around $0$ rad/s, the post-fault marginals show larger spread or dispersion of the probability mass, as expected. {\red{That the generator 4 states in Case II have larger uncertainties than in Case I, is also visible in the bivariate marginal plots if we compare the fourth row in Fig. \ref{fig:IEEE14marginalsCase0} with the same in Fig. \ref{fig:IEEE14marginalsCase1}.}}

@@

\begin{table}[t]
\centering
\begin{tabular}{| l | l |} 
\hline
Parameter description & Values\\
\hline\hline
&\\
nominal frequency & $f_{0} = 60$ Hz\\
&\\
inertia & $m_{i}\in{\rm{Unif}}\left([2,12]\right)/2\pi f_{0}$\\
& \\
damping coefficient & $\gamma_{i}\in{\rm{Unif}}\left([20,30]\right)/2\pi f_{0}$\\
& \\
diffusion coefficient & $\sigma_{i}\in{\rm{Unif}}\left([1,5]\right)$\\
& \\
tangent of phase shift & $\tan\varphi_{ij}\begin{cases}
=0 & \text{for}\; i=j\\
\in{\rm{Unif}}\left([0,0.25]\right)& \text{for}\; i\neq j
\end{cases}$\\
& \\
effective power input & $P_{i}\in{\rm{Unif}}\left([0,10]\right)$\\
& \\
coupling coefficient & $k_{ij}\begin{cases}
=0 & \text{for}\; i=j\\
\in{\rm{Unif}}\left([0.7,1.2]\right) & \text{for}\; i\neq j	
\end{cases}$\\
& \\
\hline
\end{tabular}
\caption{Parameters used for the numerical simulation in Sec. \ref{subsec:NumSimSynthetic}. The indices $i,j\in\{1,\hdots,n\}$ where the number of generators $n=50$ in Sec. \ref{subsec:NumSimSynthetic}.}
\label{table:ParamUncertaintiesSynthetic}
\vspace*{-0.1in}
\end{table}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.95\linewidth]{ComputationalTimeSynthetic50Gen.png}
\caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{50}\times\mathbb{R}^{50}$ for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
\vspace*{-0.1in}
\label{fig:CompTimeSynthetic}
\end{figure}
\begin{figure}[t]
\centering
\includegraphics[width=0.95\linewidth]{RelativeErrorMeanMCVersusProxSynthetic50Gen.png}
\caption{\small{The relative error between the empirical mean (i.e., Monte Carlo estimate) $\bm{\mu}_{k}^{\text{MC}}\in\mathbb{T}^{50}\times\mathbb{R}^{50}$, and the mean $\bm{\mu}_{k}^{\text{Prox}}\in\mathbb{T}^{50}\times\mathbb{R}^{50}$ obtained using the proximal updates of the joint state PDFs, for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
\vspace*{-0.1in}
\label{fig:RelErrSynthetic}
\end{figure}

@@

\section{Conclusion}\label{sec:conclusion}
We proposed a nonparametric algorithm to propagate the joint state PDFs subject to networked power system dynamics with stochastic initial conditions, parameters and process noise. Our development is built on recent advances in generalized gradient flows on the space of PDFs, and new analytical results specific to the power system's stochastic dynamics that are presented in this paper. The novelty of the proposed proximal algorithm is that it does not approximate the dynamical or statistical nonlinearities. Instead, the algorithm allows gridless computation by actually exploiting the structure of the PDE governing the evolution of the joint state PDF that is induced by the underlying nonlinear sample path dynamics of a networked power system. Numerical case studies reveal that the proposed computational framework can be a scalable way to propagate stochastic uncertainties in realistic power networks.

>>>>

\begin{figure}[t]
\centering
\includegraphics[width=.7\linewidth]{BlockDiagm.pdf}
\caption{\small{Schematic of the proposed proximal algorithm for propagating the joint state PDF as probability weighted scattered point cloud $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$. The states $\{\tilde{\bm{x}}_{k}^{i}\}_{i=1}^{N}$ are updated by the Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}); the corresponding probability weights $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are updated via discrete version of the proximal recursion (\ref{KineticProx}) as detailed in Sec. \ref{sec:ProxAlgo}.}}
\label{fig:BlockDiagm}
\vspace*{-0.2in}
\end{figure}

@@

\begin{remark}\label{remark:timevaryingEffectivePowerInput}
The proposed computational framework is also applicable when $P_{i}^{\rm{mech}}, P_{i}^{\rm{load}}$, and thus $P_{i}$ in \eqref{defPi}, are bounded time-varying functions. This can be the case, for instance, with rapid fluctuations from renewables in the short term.
\end{remark}

@@

\begin{figure}[htpb]
\centering
\includegraphics[width=0.95\linewidth]{ComputationalTimeSyntheticIEEE14busCase1and2combined.png}
\caption{\small{The computational times for propagating the transient joint state PDFs over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ for the simulation set up in Sec. \ref{subsec:NumSimIEEEBus}.}}
\vspace*{-0.1in}
\label{fig:CompTimeIEEE14bus}
\end{figure}

@@

\noindent\textbf{Case I.} \textbf{Nominal case} where the $P_{i}^{\text{mech}}$ and $P_{i}^{\text{load}}$ are from the steady state power flow solutions without the process noise, and assuming to remain constant over the simulation time horizon. The transient stochastic simulations (i.e., with process noise) are then performed using the proposed framework.

\noindent\textbf{Case II.} \textbf{Post-contingency case} where the line 13 (see Fig. \ref{fig:Motivating}) fails at time $t=0$, and starting with the parameters and initial conditions of \textbf{Case I}, we simulate the post-contingency stochastic state evolution with process noise. %In other words, \textbf{Case II} starts  and tracks the stochastic states towards a post-contingency equilibrium.

@@

For both the cases mentioned above, we used the randomly generated noise coefficients $\sigma_{i}\in [1,5]$ as
\[(\sigma_{1},\hdots,\sigma_5) = \left(2.4628, 4.9266, 4.8724, 1.4215, 3.8681\right).\]

To account the pertinent geometry of $\mathbb{T}^{n}\times\mathbb{R}^{n}$, we take the initial joint PDF as
\begin{align}
\rho_{0}&\equiv\rho(t=0,\bm{\theta}(0),\bm{\omega}(0)) = \vartheta_{0}\left(\bm{\theta}(0)\right)\Omega_{0}\left(\bm{\omega}(0)\right)\nonumber\\
&=\underbrace{\left(\prod_{i=1}^{n=5}\dfrac{\exp\left(\kappa_{i}\cos\left(2\theta_{i}(0)-\mu_{\theta_{i}(0)}\right)\right)}{2\pi I_{0}(\kappa_{i})}\right)}_{=:\vartheta_{0}\left(\bm{\theta}(0)\right)}\nonumber\\
	&\qquad\times \underbrace{{\rm{Unif}}\left([-0.1,0.1]\;\text{rad/s}\right)}_{=:\Omega_{0}\left(\bm{\omega}(0)\right)},
\label{rho0IEEE14}		
\end{align}
whose $\bm{\theta}$ marginal $ \vartheta_{0}\left(\bm{\theta}(0)\right)$ is a product von Mises PDF\footnote{the factor $2$ inside the cosine ensures the range $[0,2\pi)$; it can be dispensed if we instead use the range $[-\pi,\pi)$ for the angular variables, as common in the directional statistics literature \cite{mardia2009directional}.} \cite{mardia2008multivariate,mardia2014some} supported over $\mathbb{T}^{5}$ with mean angles (in rad)
\begin{align}
\left(\mu_{\theta_{1}}(0),\hdots,\mu_{\theta_{5}}(0)\right)=\left(0, 6.1963, 6.0612, 6.0350, 6.0500\right),
\label{MeanAngleInitialIEEE14}	
\end{align}
and concentration parameter $\kappa_{i}\geq 0$ set as 
\[(\kappa_{1},\hdots,\kappa_5) = \left(5,6,7,4,5\right).\]
In \eqref{rho0IEEE14}, $I_{0}(\cdot)$ denotes the modified Bessel function of the first kind with order zero, given by $I_{0}(\kappa) = \sum_{r=0}^{\infty} \frac{\left(\kappa^{2}/4\right)^{r}}{(r!)^{2}}$, and can be evaluated via MATLAB command \texttt{besseli(0,$\cdot$)}.

@@

Fig. \ref{fig:IEEE14omegamarginalsCase12} shows the corresponding univariate $\omega_{i}$-marginal PDFs. In particular, Fig. \ref{fig:IEEE14omegamarginalsCase12}(c) helps compare the $\omega_{4}$ marginal evolution for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios. Since Case II considers the failure of line 13, from Fig. \ref{fig:Motivating} we intuitively expect that the transient effects of this fault will be prominent for bus 6 (generator 4). Indeed, Fig. \ref{fig:IEEE14omegamarginalsCase12}(c) shows that although the $\omega_{4}$ marginals in both Case I and II, at all times have highest probability around $0$ rad/s, the post-fault marginals show larger spread or dispersion of the probability mass, as expected. That the generator 4 states in Case II have larger uncertainties than in Case I, is also visible in the bivariate marginal plots if we compare the fourth row in Fig. \ref{fig:IEEE14marginalsCase0} with the same in Fig. \ref{fig:IEEE14marginalsCase1}.

@@

\begin{table}[t]
\centering
\begin{tabular}{| l | l |} 
\hline
Parameter description & Values\\
\hline\hline
&\\
nominal frequency & $f_{0} = 60$ Hz\\
&\\
inertia & $m_{i}\in{\rm{Unif}}\left([2,12]\right)/2\pi f_{0}$\\
& \\
damping coefficient & $\gamma_{i}\in{\rm{Unif}}\left([20,30]\right)/2\pi f_{0}$\\
& \\
diffusion coefficient & $\sigma_{i}\in{\rm{Unif}}\left([1,5]\right)$\\
& \\
tangent of phase shift & $\tan\varphi_{ij}\begin{cases}
=0 & \text{for}\; i=j\\
\in{\rm{Unif}}\left([0,0.25]\right)& \text{for}\; i\neq j
\end{cases}$\\
& \\
effective power input & $P_{i}\in{\rm{Unif}}\left([0,10]\right)$\\
& \\
coupling coefficient & $k_{ij}\begin{cases}
=0 & \text{for}\; i=j\\
\in{\rm{Unif}}\left([0.7,1.2]\right) & \text{for}\; i\neq j	
\end{cases}$\\
& \\
\hline
\end{tabular}
\caption{Parameters used for the numerical simulation in Sec. \ref{subsec:NumSimSynthetic}. The indices $i,j\in\{1,\hdots,n\}$ where the number of generators $n=50$ in Sec. \ref{subsec:NumSimSynthetic}.}
\label{table:ParamUncertaintiesSynthetic}
\vspace*{-0.35in}
\end{table}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.95\linewidth]{ComputationalTimeSynthetic50Gen.png}
\caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{50}\times\mathbb{R}^{50}$ for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
\vspace*{-0.1in}
\label{fig:CompTimeSynthetic}
\end{figure}
\begin{figure}[t]
\centering
\includegraphics[width=0.95\linewidth]{RelativeErrorMeanMCVersusProxSynthetic50Gen.png}
\caption{\small{The relative error between the empirical mean (i.e., Monte Carlo estimate) $\bm{\mu}_{k}^{\text{MC}}\in\mathbb{T}^{50}\times\mathbb{R}^{50}$, and the mean $\bm{\mu}_{k}^{\text{Prox}}\in\mathbb{T}^{50}\times\mathbb{R}^{50}$ obtained using the proximal updates of the joint state PDFs, for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
\vspace*{-0.2in}
\label{fig:RelErrSynthetic}
\end{figure}

@@

\section{Conclusion}\label{sec:conclusion}
We proposed a nonparametric algorithm to propagate the joint state PDFs subject to networked power system dynamics with stochastic initial conditions, parameters and process noise. Our development is built on recent advances in generalized gradient flows on the space of PDFs, and new analytical results specific to the power system's stochastic dynamics that are presented in this paper. The novelty of the proposed proximal algorithm is that it does not approximate the dynamical or statistical nonlinearities. Instead, the algorithm allows gridless computation by actually exploiting the structure of the PDE governing the evolution of the joint state PDF that is induced by the underlying nonlinear sample path dynamics of a networked power system. Numerical case studies reveal that the computational framework can be a scalable way to propagate stochastic uncertainties in realistic power networks.

### 2021-08-28 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_10pg.tex
updates

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{UncPropBlockDiagm.pdf}
\caption{\small{Block diagram for joint state PDF propagation.}}
\label{fig:UncertaintyPropagation}
\vspace*{-0.1in}
\end{figure}

@@

\begin{figure}[tb]
\centering
\includegraphics[width=0.9\linewidth]{ProxSchmatic.pdf}
\caption{\small{A schematic of approximating the flow of the joint PDF trajectory $\rho(t,\bm{x})$ via the sequence $\{\varrho_{k}\}_{k\in\mathbb{N}}$ generated by a variational recursion over $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$. The curve $\rho(t,\bm{x})$ solves a kinetic Fokker-Planck PDE initial value problem. Points on this curve (shown as three filled circled markers at three specific instances $t_0,t_1,t_2$) are joint PDFs. While the joint PDFs are supported over the finite dimensional base manifold $\mathbb{T}^{n}\times\mathbb{R}^{n}$, the proximal updates $\{\varrho_{k}\}_{k\in\mathbb{N}}$ (shown as circled markers with no face-color) evolve over the infinite dimensional manifold $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$.}}
\vspace*{-0.1in}
\label{fig:ProxSchematic}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=.7\linewidth]{BlockDiagm.pdf}
\caption{\small{Schematic of the proposed proximal algorithm for propagating the joint state PDF as probability weighted scattered point cloud $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$. The states $\{\tilde{\bm{x}}_{k}^{i}\}_{i=1}^{N}$ are updated by the Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}); the corresponding probability weights $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are updated via discrete version of the proximal recursion (\ref{KineticProx}) as detailed in Sec. \ref{sec:ProxAlgo}.}}
\label{fig:BlockDiagm}
\vspace*{-0.1in}
\end{figure}

@@

\begin{remark}\label{remark:timevaryingEffectivePowerInput}
The proposed computational framework can be applied as is when $P_{i}^{\rm{mech}}, P_{i}^{\rm{load}}$, and thus $P_{i}$ in \eqref{defPi}, are bounded time-varying functions. {\red{This can be the case, for instance, when ...}} 	
\end{remark}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=0.9\linewidth]{IEEE14marginalsCase1.png}
\caption{\small{Time evolution of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs of the generator $i=1,\hdots,5$, for the Kron-reduced IEEE 14 bus simulation setup \textbf{Case II} described in Sec. \ref{subsec:NumSimIEEEBus}. The $5$ rows above correspond to the $5$ generator nodes; these are buses 1, 2, 3, 6 and 8 in Fig. \ref{fig:Motivating}. The columns correspond to the time snapshots. The colorbar at the bottom shows the marginal PDF values. The $\theta_i$ are in rad, the $\omega_i$ are in rad/s.}}
\vspace*{-0.1in}
\label{fig:IEEE14marginalsCase1}
\end{figure*}

@@

\noindent\textbf{Case I.} {\red{Nominal case where ...}}

\noindent\textbf{Case II.} {\red{post-contingency simulation after the failure of line 13 (see Fig. \ref{fig:Motivating}).}}

@@

For all the cases mentioned above, we used the randomly generated noise coefficients $\sigma_{i}\in [1,5]$ as
\[(\sigma_{1},\hdots,\sigma_5) = \left(2.4628, 4.9266, 4.8724, 1.4215, 3.8681\right).\]

@@

Fig. \ref{fig:IEEE14marginalsCase0} and Fig. \ref{fig:IEEE14marginalsCase1} show the time snapshots of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs for \textbf{Case I} and \textbf{Case II} respectively, associated with the generator $i=1,\hdots, n=5$. These marginals are computed using the corresponding weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.

 %Notice that in Fig. \ref{fig:IEEE14marginals}, the horizontal axes of the subfigures have periodic boundary conditions, i.e., $(\theta_{i},\omega_{i})\in\mathbb{T}\times\mathbb{R}$ (cylinder).

Fig. \ref{fig:IEEE14omegamarginalsCase12} shows the corresponding univariate $\omega_{i}$-marginal PDFs. In particular, Fig. \ref{fig:IEEE14omegamarginalsCase12}(c) helps compare the $\omega_{4}$ marginal evolution for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios. Since Case II considers the failure of line 13, from Fig. \ref{fig:Motivating} we intuitively expect that the transient effects of this fault will be prominent for bus 6 (generator 4). Indeed, Fig. \ref{fig:IEEE14omegamarginalsCase12}(c) shows that although the $\omega_{4}$ marginals in both Case I and II, at all times have highest probability around $0$ rad/s, the post-fault marginals show larger spread or dispersion of the probability mass, as expected. {\red{That the generator 4 states in Case II have larger uncertainties than in Case I, is also visible in the bivariate marginal plots if we compare the fourth row in Fig. \ref{fig:IEEE14marginalsCase0} with the same in Fig. \ref{fig:IEEE14marginalsCase1}.}}

@@

>>>>

\begin{figure}
\centering
\includegraphics[width=0.8\linewidth]{UncPropBlockDiagm.pdf}
\caption{\small{Block diagram for joint state PDF propagation.}}
\label{fig:UncertaintyPropagation}
\vspace*{-0.15in}
\end{figure}

@@

\begin{figure}[tb]
\centering
\includegraphics[width=0.7\linewidth]{ProxSchmatic.pdf}
\caption{\small{A schematic of approximating the flow of the joint PDF trajectory $\rho(t,\bm{x})$ via the sequence $\{\varrho_{k}\}_{k\in\mathbb{N}}$ generated by a variational recursion over $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$. The curve $\rho(t,\bm{x})$ solves a kinetic Fokker-Planck PDE initial value problem. Points on this curve (shown as three filled circled markers at three specific instances $t_0,t_1,t_2$) are joint PDFs. While the joint PDFs are supported over the finite dimensional base manifold $\mathbb{T}^{n}\times\mathbb{R}^{n}$, the proximal updates $\{\varrho_{k}\}_{k\in\mathbb{N}}$ (shown as circled markers with no face-color) evolve over the infinite dimensional manifold $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$.}}
\vspace*{-0.1in}
\label{fig:ProxSchematic}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=.6\linewidth]{BlockDiagm.pdf}
\caption{\small{Schematic of the proposed proximal algorithm for propagating the joint state PDF as probability weighted scattered point cloud $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$. The states $\{\tilde{\bm{x}}_{k}^{i}\}_{i=1}^{N}$ are updated by the Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}); the corresponding probability weights $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are updated via discrete version of the proximal recursion (\ref{KineticProx}) as detailed in Sec. \ref{sec:ProxAlgo}.}}
\label{fig:BlockDiagm}
\vspace*{-0.2in}
\end{figure}

@@

\begin{remark}\label{remark:timevaryingEffectivePowerInput}
The proposed computational framework is also applicable when $P_{i}^{\rm{mech}}, P_{i}^{\rm{load}}$, and thus $P_{i}$ in \eqref{defPi}, are bounded time-varying functions. This can be the case, for instance, with rapid fluctuations from renewables in the short term.
\end{remark}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=0.9\linewidth]{IEEE14BivariateMarginalComparison.pdf}
\caption{\small{Comparison of the $(\theta_{4},\omega_{4})$ marginals for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios in the IEEE 14 bus simulation described in Sec. \ref{subsec:NumSimIEEEBus}.}}
\vspace*{-0.1in}
\label{fig:IEEE14marginalsBivariateComparison}
\end{figure*}

@@

\noindent\textbf{Case I.} \textbf{Nominal case} where the $P_{i}^{\text{mech}}$ and $P_{i}^{\text{load}}$ are from the steady state power flow solutions without the process noise, and assuming to remain constant over the simulation time horizon. The transient stochastic simulations (i.e., with process noise) are then performed using the proposed framework.

\noindent\textbf{Case II.} \textbf{Post-contingency case} where the line 13 (see Fig. \ref{fig:Motivating}) fails at time $t=0$, and starting with the parameters and initial conditions of \textbf{Case I}, we simulate the post-contingency stochastic state evolution with process noise.

@@

For both the cases mentioned above, we used the randomly generated noise coefficients $\sigma_{i}\in [1,5]$ as
\[(\sigma_{1},\hdots,\sigma_5) = \left(2.4628, 4.9266, 4.8724, 1.4215, 3.8681\right).\]

@@

Fig. \ref{fig:IEEE14marginalsCase0} shows the time snapshots of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs for \textbf{Case I} associated with the generator $i=1,\hdots, n=5$. These marginals are computed using the corresponding weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.

Since \textbf{Case II} considers the failure of line 13, from Fig. \ref{fig:Motivating} we intuitively expect that the transient effects of this fault will be prominent for bus 6 (generator 4). The comparison of the $(\theta_4,\omega_4)$ bivariate marginals between \textbf{Cases I} and \textbf{II} shown in Fig. \ref{fig:IEEE14marginalsBivariateComparison} indeed confirms that the post contingency transient stochastic states for bus 6 have higher dispersion or spread of the probability mass.

Likewise, Fig. \ref{fig:IEEE14omegamarginalsCase12} compares the $\omega_{4}$ marginal evolution for bus 6 in the nominal (\textbf{Case I}) and post-fault (\textbf{Case II}) scenarios. As expected, Fig. \ref{fig:IEEE14omegamarginalsCase12} highlights that although the $\omega_{4}$ marginals in both \textbf{Cases I} and \textbf{II}, at all times have highest probability around $0$ rad/s, the post-fault marginals show larger dispersion. %{\red{That the generator 4 states in Case II have larger uncertainties than in Case I, is also visible in the bivariate marginal plots if we compare the fourth row in Fig. \ref{fig:IEEE14marginalsCase0} with the same in Fig. \ref{fig:IEEE14marginalsCase1}.}}

@@

% expanded version of this paper (exceeds 10 pages)

\bibitem{halder2021stochasticExpanded}
A.~Halder, K.F.~Caluya, P.~Ojaghi, and X.~Geng, ``Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions", \emph{extended preprint}, \\
doi: \url{http://dx.doi.org/10.13140/RG.2.2.30956.28805}, 2021.
